---
title: "R Notebook"
output: html_notebook
---

http://www.math.uah.edu/stat/data/Challenger2.txt
# 23 previous space shuttle launches before the Challenger disaster
# T is the temperature in Fahrenheit, I is the O-ring damage index

```{r}
oring=read.table("http://www.math.uah.edu/stat/data/Challenger2.txt",header=T)
attach(oring)
#note: masking T=TRUE
plot(T,I)
```

```{r}
oring.lm=lm(I~T)
summary(oring.lm)
```

# add fitted line to scatterplot
```{r}
plot(T,I)
lines(T,fitted(oring.lm))            
```
# 95% posterior interval for the slope
```{r}
-0.24337 - 0.06349*qt(.975,21)
-0.24337 + 0.06349*qt(.975,21)
```
# note that these are the same as the frequentist confidence intervals

# the Challenger launch was at 31 degrees Fahrenheit
# how much o-ring damage would we predict?
# y-hat
```{r}
18.36508-0.24337*31
coef(oring.lm)
coef(oring.lm)[1] + coef(oring.lm)[2]*31  
```

# posterior prediction interval (same as frequentist)
```{r}
predict(oring.lm,data.frame(T=31),interval="predict")  
10.82052-2.102*qt(.975,21)*sqrt(1+1/23+((31-mean(T))^2/22/var(T)))
```

# posterior probability that damage index is greater than zero
1-pt((0-10.82052)/(2.102*sqrt(1+1/23+((31-mean(T))^2/22/var(T)))),21)


http://www.math.uah.edu/stat/data/Galton.txt
# Galton's seminal data on predicting the height of children from the 
# heights of the parents, all in inches

```{r}
heights=read.table("http://www.math.uah.edu/stat/data/Galton.txt",header=T)
attach(heights)
names(heights)
```

```{r}
pairs(heights)
summary(lm(Height~Father+Mother+Gender+Kids))
summary(lm(Height~Father+Mother+Gender))
heights.lm=lm(Height~Father+Mother+Gender)
```

# each extra inch taller a father is is correlated with 0.4 inch extra
  height in the child
# each extra inch taller a mother is is correlated with 0.3 inch extra
  height in the child
# a male child is on average 5.2 inches taller than a female child
# 95% posterior interval for the the difference in height by gender
5.226 - 0.144*qt(.975,894)
5.226 + 0.144*qt(.975,894)

# posterior prediction interval (same as frequentist)
predict(heights.lm,data.frame(Father=68,Mother=64,Gender="M"),interval="predict")
predict(heights.lm,data.frame(Father=68,Mother=64,Gender="F"),interval="predict")

# Quiz

```{r}
dat=read.table("http://www.stat.ufl.edu/~winner/data/pgalpga2008.dat",col.names=c('dist', 'accuracy', 'sex'))
datF <- subset(dat, sex==1, select=1:2)
datM <- subset(dat, sex==2, select=1:2)
dat
```

```{r}
plot(datF$dist, datF$accuracy)
plot(datM$dist, datM$accuracy)
```

```{r}
datF.lm=lm(datF$accuracy~datF$dist)
summary(datF.lm)
```

```{r}
coef(datF.lm)[1] + coef(datF.lm)[2]*260
```

```{r}
predict(datF.lm,data.frame(T=31),interval="predict")  
10.82052-2.102*qt(.975,21)*sqrt(1+1/23+((31-mean(T))^2/22/var(T)))
```

## multiple regression
```{r}
dat.lm=lm(dat$accuracy~dat$dist+dat$sex)
summary(dat.lm)
```

```{r}
plot(fitted(dat.lm), residuals(dat.lm))
```